---
title: "2darbas2"
author: "Asta Kvedaraite & Marius Survila"
date: "April 1, 2020"
output: html_document
---

```{r}
library(minfi)
inp <- dirname(getwd())
path<- file.path(inp,"output", "RGset.rds")

target <- readRDS(path)

pd <- pData(target)
as.data.frame(pd[2:4,])
rm(pd)
```

#7. p-val > 0.01 => unreliable\
Removes samples that have > 1%  unreliable values\
(count unreliable values / count all values > 0.01)
```{r}

if (require(minfi)) {
  failed <- detectionP(target) > 0.01
  #2 - column, 1- row in a matrix
  ind <- which(apply(failed, 2, function(x) (count(x)/length(x))) > 0.01)
  ind <- as.numeric(gsub(".* ", "",ind))
}
target <- target[,-ind]
dim(target)
rm(ind)
```

#8.Normalisation with peprocessSWAN - \
Subset-Quantile Within Array Normalisation \
For Illumina Infinium HumanMethylation450 BeadChips.\
It allows Infinium I and II type probes on a single\
array to be normalized together.\

Also deleting all rows with more than 1% unreliable data
```{r}
MSet.swan <- preprocessSWAN(target)

if (require(minfi)) {
   #2 - column, 1- row in a matrix
  ind <- which(apply(failed, 1, function(x) (count(x)/length(x))) > 0.01)
  ind <- as.numeric(gsub(".* ", "",ind))
}
MSet.swan <- MSet.swan[-ind,]
dim(MSet.swan)

rm(target)
rm(ind)
rm(failed)
```

#10. Drop positions with "CH" and \
#SNP - positions that are near DNR polymorphism\
(dropRS)

```{r}
MSet.swan <- dropMethylationLoci(MSet.swan, dropRS = TRUE, dropCH = TRUE)
dim(MSet.swan)

```

# 11 Saving beta, pData and annotation to rds
```{r}
beta <- getBeta(MSet.swan)
dim(beta)

path<- file.path(inp,"output", "beta.rds")
saveRDS(beta, path)

rm(beta)

data <- pData(MSet.swan)

path<- file.path(inp,"output", "pData.rds")
saveRDS(data, path)

rm(data)

annotation <- getAnnotation(MSet.swan)

path<- file.path(inp,"output", "annotation.rds")
saveRDS(annotation, path)

rm(annotation)

```