---
title: "2darbas"
author: "Asta Kvedaraite & Marius Survila"
date: "March 26, 2020"
output: html_document
---
```{r}
#Install minfiData package if needed

#  if (!requireNamespace("BiocManager", quietly = TRUE))
#      install.packages("BiocManager")
#  BiocManager::install("minfiData")

```

```{r}

dir <- dirname(getwd())
baseDir<- file.path(dir,"input","GSE125895","supplement", "GSE125895_idat","/")

```
#Read samplekey data and remove unnecessary columns
```{r}
inp <- file.path(dir, "input", "GSE125895_samplekey.csv")
data <- read.csv(inp, as.is = TRUE)
rm(inp)

#Remove unnecessary columns
data <- data[-c(3:5,7,24,26,28,29,32:34,36:46,48,49)]

head(data)
```
#Renaming columns to more obvious names and simplifying data
```{r}
names(data)[names(data) == "title"] <- "sample_name"
data$sample_name <- gsub("sample_plate:", "", data$sample_name)

names(data)[names(data) == "geo_accession"] <- "slide"


names(data)[names(data) == "characteristics_ch1"] <- "sample_plate"
data$sample_plate <- gsub("sample_plate:", "", data$sample_plate)

names(data)[names(data) == "characteristics_ch1.2"] <- "brNum"
data$brNum <- gsub("brnum: ", "", data$brNum)

names(data)[names(data) == "characteristics_ch1.3"] <- "tissue"
data$tissue <- gsub("brnum: ", "", data$tissue)

names(data)[names(data) == "characteristics_ch1.4"] <- "tissue_region"
data$tissue_region <- gsub("tissue region: ", "", data$tissue_region)

names(data)[names(data) == "characteristics_ch1.5"] <- "sample"
data$sample <- gsub("disease state \\(dx\\): ", "", data$sample)


names(data)[names(data) == "characteristics_ch1.6"] <- "age"
data$age <- gsub("age: ", "", data$age)
data$age <- as.integer(as.numeric(data$age))

names(data)[names(data) == "characteristics_ch1.7"] <- "sex"
data$sex <- gsub("Sex: ", "", data$sex)

names(data)[names(data) == "characteristics_ch1.8"] <- "race"
data$race <- gsub("race: ", "", data$race)

names(data)[names(data) == "characteristics_ch1.9"] <- "negcontrol_pc1"
data$negcontrol_pc1 <- gsub("negcontrol_pc1: ", "", data$negcontrol_pc1)

names(data)[names(data) == "characteristics_ch1.10"] <- "negcontrol_pc2"
data$negcontrol_pc2 <- gsub("negcontrol_pc2: ", "", data$negcontrol_pc2)

names(data)[names(data) == "characteristics_ch1.11"] <- "neun_pos"
data$neun_pos <- gsub("neun_pos: ", "", data$neun_pos)

names(data)[names(data) == "characteristics_ch1.12"] <- "snppc1"
data$snppc1 <- gsub("snppc1: ", "", data$snppc1)

names(data)[names(data) == "characteristics_ch1.1"] <- "Array"
data$Array <- gsub("sentrix_position: ", "", data$Array)


names(data)[names(data) == "supplementary_file"] <- "Basename"
data$Basename <- gsub("ftp://ftp.ncbi.nlm.nih.gov/geo/samples/GSM3584nnn/.*/suppl/", baseDir, data$Basename)
data$Basename <- gsub("_Red.idat.gz|_Grn.idat.gz", "", data$Basename)
#head(data$Basename)
head(data)

```
#Reading idat files and creating RGChannelSet
```{r, echo = FALSE}
library(minfi)
```

```{r}
RGset <- read.metharray.exp(targets = data)

output<- file.path(dir, "output", "RGset.rds")
saveRDS(RGset, file = output)
rm(data)
rm(output)
```
#check for column count\
There are `r nrow(RGset)` rows - methylated estimates\
and `r ncol(RGset)` columns - samples\
Each column has more column data (sample_name ,age, gender, etc)
```{r}
RGset
```
#Shows several samples
```{r}

#show several rows 
pd <- pData(RGset)

#several samples
as.data.frame(pd[2:4, c(1, 2, 7, 8)])
rm(pd)

```

#Shows genomic position
```{r}
manifest <- getManifest(RGset)
manifest
as.data.frame(head(getProbeInfo(manifest)))
rm(manifest)
```


